Figures for Paper

### 2021-02-11 paper_figures.Rmd
making rmd pretty

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel")
```

@@

#### TARDBP RNA and UNC13A Cryptic
```{r}

@@

#### UNC13A RNA and UNC13A Cryptic####

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
           cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
      scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
#### UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```
#### TARDBP RNA and UNC13A IR

@@

#### barplot UNC13A relative RNA across experiments

@@

#### scatterplot UNC13BIR and UNC13B NMD

@@

#### barplot protein abundances iPSC KD

@@

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
# Detection of UNC13A
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

@@

# UNC13A detection rate by tissue

@@

# Detection UNC13A by tissue and Genotype

@@

# Junction type expression by tissue - shorter exon only collapsed

```{r}
boxplot_table %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = inclusion_reads * 10^6,
               fill ="red")) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") 
```

# Junction type expression by tissue - the longer acceptor

@@

@@

# UNC13A TPM by detection status and tissue

@@

clean_data_table %>% 
  filter(UNC13A_annotated_leaf > 10) %>% 
  mutate(unc13_d = ifelse(unc13a_cryptic_leaf_psi > 0,"detected","not-detected")) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>%  
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
     filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                "Frontal Cortex",
                                "Lumbar Spinal Cord",
                                "Temporal Cortex",
                                "Motor Cortex")) %>% 
    ggbarplot(,
              x = "detection_name",
              add = c("mean_se","jitter"),
              y = "UNC13A_TPM",
              fill = 'unc13_d',
              color = 'unc13_d',
              facet.by = 'tissue_clean',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  ggtitle("UNC13A Cryptic Detected - all tissue") +
  stat_compare_means(aes(x  = detection_name, fill = unc13_d,y = UNC13A_TPM))
```

# Correlation between STMN2 Cryptic PSI and UNC13A TPM

Separated by tissue only in ALS/FTLD-TDP
```{r}

@@

No Tissue separation

# Correlation between UNC13A CE PSI and UNC13A TPM
## tissue separated

@@

## not separated -

@@

# STMN2 and UNC13A CE PSI in tissue

```{r}
clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>% 
  mutate(unc13a_normalized_psi_to_stmn2 = log2((unc13a_cryptic_leaf_psi_full +1)/ (stmn_2_cryptic_psi_leaf + 1))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(tissue_clean) %>% 
  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
  ungroup() %>% 
  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
  ggbarplot( x = "tissue_clean", 
               y = "unc13a_normalized_psi_to_stmn2", 
              position = position_dodge(0.8),
              add = c("mean_se","jitter"),
             facet.by = 'disease_group2') +
  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
  xlab("") +
  theme(text = element_text(size = 24)) +
  theme(legend.position = 'bottom') +
  facet_wrap(~disease_group2, scales = 'free_x')
```

# boxplot of cryptic psi in als-tdp and ftld-tdp by genotype and tissue

```{r}
clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(unc13a_cryptic_leaf_psi_full > 0 ) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  ggboxplot( x = "tissue_clean", 
               y = "unc13a_cryptic_leaf_psi", 
             fill = "call",
             facet.by = 'disease_group2') +
  xlab("") +
  geom_jitter(height = 0) +
  theme(text = element_text(size = 24)) +
  theme(legend.position = 'bottom') +
  facet_wrap(~disease_group2, scales = 'free')

@@

```{r}
cortex_names = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  .$detection_name %>% unique()

@@

# boxplot of UNC over STMN2 psi in als-tdp and ftld-tdp by genotype and tissue

```{r}
cortex_comparisons = list(c("C/C \n( 121 )", "G/G \n( 54 )"),c("G/G \n( 54 )", "C/G \n( 110 )"),c("C/C \n( 121 )", "C/G \n( 110 )"))
detected_comparisons = list(c("C/C \n( 26 )", "G/G \n( 21 )"),c("G/G \n( 21 )", "C/G \n( 26 )"),c("C/C \n( 26 )", "C/G \n( 26 )"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = detection_name)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.3, height = 0) + 
  stat_compare_means() +
  ylab("STMN2 PSI")

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
    group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
    mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.3, height = 0) + 
  stat_compare_means() +
  ylab("STMN2 PSI")

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
    group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.3, height = 0) + 
  stat_compare_means() +
  facet_wrap(~disease_group2) +
  stat_compare_means( label = "p.format") + 
  # stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif") +
  ylab("STMN2 PSI")

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
    group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.3, height = 0) + 
  stat_compare_means() +
  facet_wrap(~disease_group2) +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = cortex_comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 1,size = 1.5) +
  ggtitle("All Cortex Samples")

@@

@@

clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,STMN2_TPM,UNC13A_TPM) %>% 
  mutate(unc13a_normalized_psi_to_stmn2 = log2((UNC13A_TPM +1)/ (STMN2_TPM + 1))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(tissue_clean) %>% 
  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
  ungroup() %>% 
  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
  ggbarplot( x = "tissue_clean", 
               y = "unc13a_normalized_psi_to_stmn2", 
              position = position_dodge(0.8),
              add = c("mean_se","jitter"),
             facet.by = 'disease_group2') +
  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
  xlab("") +
  theme(text = element_text(size = 24)) +
  theme(legend.position = 'bottom') +
  facet_wrap(~disease_group2, scales = 'free_x')

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(participant_id,  sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>%  add_count(participant_id) %>% 
  filter(n == 5) %>% 
  dplyr::select(-n) %>% 
  data.table::melt() %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  ggplot(aes(x = tissue_clean, y = value, color = participant_id)) + 
  geom_point() +
  geom_line(aes(group = participant_id)) +
  theme(text = element_text(size = 24)) +
  theme(legend.position = 'none')  +
  facet_grid(~variable)
  scale_color_manual(values = c("#D55E00","#0072B2"))

@@

# Correlation UNC13A PSI and STMN2 PSI with genotype effect

>>>>

```{css, echo = FALSE}
h1, #TOC>ul>li {
  color: #000000;
    font-weight: bold;
}

h2, #TOC>ul>ul>li {
  color: #41494D;
}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
knitr::opts_chunk$set(
  comment = '', fig.width = 9, fig.height = 9
)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel")
```

@@

## TARDBP RNA and UNC13A Cryptic

@@

## UNC13A RNA and UNC13A Cryptic

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
           cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
      scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
## UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```
## TARDBP RNA and UNC13A IR

@@

## barplot UNC13A relative RNA across experiments

@@

## scatterplot UNC13BIR and UNC13B NMD

@@

# iPSC protein abundance mass spec

@@

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
## Pecent of patients UNC13A detected
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

@@

## Pecent of samples UNC13A detected by tissue

@@

## Pecent of samples UNC13A detected by tissue and genotype

@@

## UNC13A CE level across tissue

This only plots the short novel acceptor and novel donor

```{r onlyshort}
boxplot_table %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = inclusion_reads * 10^6,
               fill ="red")) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") 
```

## UNC13A CE level across tissue - split by junction type

@@

## UNC13A CE level across disease types

@@

# UNC13A RNA expression (TPM)

## UNC13A TPM by disease and tissue

```{r}
comps = unique(clean_data_table$disease_group2) %>% gsub("Control", " Control",.) %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
 filter(tissue_clean %in% 
           c("Frontal Cortex", 
             "Lumbar Spinal Cord", 
             "Cervical Spinal Cord",
             "Motor Cortex", 
             "Temporal Cortex",
             "Cerebellum") )%>% 
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none')
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM

only in ALS/FTLD-TDP

## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations

@@

@@

## Correlation between UNC13A CE PSI and UNC13A TPM

## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated

@@

# Effect of UNC13A genotype

## UNC13A psi by genotype
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

@@

## UNC13A psi by genotype - tissue split
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

@@

## UNC13A psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(call) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

@@

## STMN2 psi by genotype 
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

@@

## STMN2 psi by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

## STMN2 psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

## Correlation UNC13A PSI and STMN2 PSI with genotype effect

@@

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'call',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = call),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32))

@@

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected
